---
title: 'Part 1: Binary Logistic Regression'
author: "Anna Talken"
date: "1/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(jtools)
```


```{r}
penguins %>%
  select(species, bill_length_mm: sex) %>% 
  ggpairs(aes(color=species))
```

Create a subset with only Adelie and Chinstrap
```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>% 
  mutate(species = fct_drop(species)) #use mutate to edit existing columns, fct_drop is factor drop which is used to drop any factors where observations do not exist. In this case, we are using fct_drop to drop the "gentoo" because we want to run logistic regression only using Adelie and Chinstrap. Using mutate to override "species" with a new species column.
```

Data viz of two species
```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y=flipper_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)
```

Binary Logistic Regression

- body mass
- flipper length
- penguin sex

```{r}
# using 'glm' which is generalized linear model b/c in binary logistic regression, the right hand side of the equation is linear. designating 'family = binomial' indicates that the left hand side will be log odds and a binomial logistic regression
adelie_chinstrap_blr <- glm(species ~ body_mass_g + flipper_length_mm +sex,
                            data = adelie_chinstrap,
                            family = "binomial")
```

Take a look at the outcomes

```{r}
adelie_chinstrap_blr

summary(adelie_chinstrap_blr) #shows more info on the blr equation (shows p-value for variables)

#interpret results for lefthand side to be probability for chinstrap
#Example interpretation: positive value associated with flipper length, means that with an increase in flipper length, the odds that it is a chinstrap penguin is increasing
# Another example, if a penguin is male, it is less likely to be a chinstrap penguin
# Another example, as body mass increases, the odds of the penguin being a chinstrap might be going down (that value is non-significant...but looking at the negative value of the coefficient that is what it means)

```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```

What are the actual probabilities of the chinstrap for each of the existing observations in the adelie_chinstrap data frame?

```{r}
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict = "response")

#the .fitted column shows the model's predicted percent chance of that specific penguin being deemed as a Chisntrap penguin (we want it to be LOW for Adelies)
```


Plotting fitted probabilities
```{r}
ggplot(data = blr_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = sex, pch = species)) + #pch = point style classifier
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = "Flipper Length",
       y = "probability of outcome Chinstrap")
```
Plot model based on different variables

```{r}
effect_plot(adelie_chinstrap_blr,
            pred = flipper_length_mm,
            y.label = "Probability of Chinstrap")
#gives quick view of overall trend between flipper length and being a Chinstrap penguin

effect_plot(adelie_chinstrap_blr,
            pred = body_mass_g,
            y.label = "Probability of Chinstrap",
            interval = TRUE)
```

What's the probability that a female penguin weighing 3410 g with a flipper length of 192 mm is a Chinstrap?

```{r}
ex1 <- predict(adelie_chinstrap_blr,
               data.frame(sex = "female", 
                          body_mass_g = 3410,
                          flipper_length_mm = 192),
               type = "response") #show as a percentage

ex1
```

























